The \project{unRAVE} catalog of X.X million detailed stellar abundances for chemodynamic studies

### 2016-08-16 article/unrave.tex
Minor edits

\begin{abstract}
On 14 September 2016 the \project{Gaia} mission will release positions, proper 
motions, and parallaxes for about two million stars.  Most studies intending to
make use of those data would benefit enormously by coupling that astrometry with
stellar chemical abundances.  The RAdial Velocity Experiment (\project{RAVE}) 
survey has acquired spectra for approximately 290,000 stars in common with the 
first \project{Gaia} data release, constituting the largest overlap of any 
stellar spectroscopic sample.  Here we perform an independent analysis of the 
\project{RAVE} spectra using a new non-parametric implementation of \thecannon\ 
that employs prior probabilities on atomic line formation and stellar 
astrophysical parameters.  We deliver effective temperature $\teff$, surface 
gravity $\logg$, and chemical abundances of up to seven elements (Mg, Al, Si, 
Ca, Ti, Fe, Ni) for \Nstars\ stars.  We validate our results internally from 
repeat visits, and externally against existing high-resolution spectroscopic 
surveys.  The typical precision in chemical abundances is X.XX~dex.  
% For stars not in TGAS, we calculate distances
\end{abstract}

@@

% These scientific questions were only some of them that prompted Gaia.
% What will Gaia provide eventually.
% In the first data release what do we get.
% While this is an order of magnitude better than Hipparcos, if we want to use these data to understand the chemical evolution of the galaxy then we need additional information.
% until abundances are available from gaia RVS, we rely on ground-based spectroscopic surveys to provide the chemistry

% Gaia will produce a very clear snapshot, producing astrometry space motions

% but if we want to infer the past then we need additional information about how the galaxy has been enriched. so we need chemical abundances.
% the number of stars that overlap with Gaia DR1 between stellar surveys is very different, due to targeting, etc
% and those stars are not on the same metallicity scale
% therefore a large-scale catalog of homogenous abundances that overlap with Gaia DR1 is currently lacking

\stub{Chemistry and kinematics are essential for unfolding the Galaxy, etc.}

\stub{
Gaia will do this, and until Gaia releases abundances for stars from RVS 
spectra, we rely on ground-based surveys to provide chemical abundances.
This includes GES, APOGEE, etc.
}

The first \project{Gaia} data release includes positions, proper motions, and 
parallaxes for approximately two million stars in the Tycho-2 catalog \citep{Hog_2000}.  
After cross-matching all major stellar spectroscopic surveys\footnote{Specifically 
we cross-matched the Tycho-2 catalog against {APOGEE} \citep{Zasowski_2013}, 
\project{GALAH} \citep{DeSilva_2015}, \project{Gaia-ESO} \citep{Gilmore_2012,Randich_2013}, \project{LAMOST}
\citep{Cui_2012}, and \project{RAVE} \citep{Steinmetz_2006}}, we found the RAdial Velocity
Experiment (hereafter \rave) 
survey to have the largest overlap with the first \project{Gaia} data release: 
292,036 stars.  We then used the \project{Gaia} universe model \citep{Robin_2012} to 
estimate the precision in parallax and proper motions that will be available in 
\project{Gaia} DR1 for every star in these cross-matches.  Comparing the 
expected precision to what is currently available, we further found that the 
\project{RAVE} survey will benefit most from \project{Gaia} DR1.  The distances
of 63\% of the \project{RAVE}--\project{Gaia} DR1 overlap sample (182,862 stars)
are expected to improve with the first \project{Gaia} data release, and 47\% of
stars likely to have better proper motions (137,211 stars).  This motivated us to
examine what chemical abundance information is currently available for \rave, and
to evaluate whether we could contribute to this existing set.

@@

>>>>

\begin{abstract}
On 14 September 2016 the \project{Gaia} mission will release positions, proper 
motions, and parallaxes for about two million stars.  Most studies intending to
make use of those data would benefit enormously by coupling that astrometry with
stellar chemical abundances.  The RAdial Velocity Experiment (\project{RAVE}) 
survey has acquired spectra for 292,036 stars in common with the 
first \project{Gaia} data release, constituting the largest overlap of any 
stellar spectroscopic sample.  Here we perform an independent analysis of the 
\project{RAVE} spectra using a new non-parametric implementation of \thecannon\ 
that employs prior probabilities on atomic line formation and stellar 
astrophysical parameters.  We deliver effective temperature $\teff$, surface 
gravity $\logg$, and chemical abundances of up to seven elements (Mg, Al, Si, 
Ca, Ti, Fe, Ni) for \Nstars\ stars.  We validate our results internally from 
repeat visits, and externally against existing high-resolution spectroscopic 
surveys.  The typical precision in chemical abundances is X.XX~dex.  
We provide improved distances by isochrone-fitting for \project{RAVE} stars
where astrometric solutions are not available in the first \project{Gaia}
data release.
\end{abstract}

@@

% Introducing Gaia
% Gaia will produce a very clear snapshot, producing astrometry space motions
% What will Gaia provide eventually.
% In the first data release what do we get.
% While this is an order of magnitude better than Hipparcos, if we want to use these data to understand the chemical evolution of the galaxy then we need additional information.
% until abundances are available from gaia RVS, we rely on ground-based spectroscopic surveys to provide the chemistry

The first \project{Gaia} data release will include positions, proper motions, and 
parallaxes for approximately two million stars in the Tycho-2 \citep{Hog_2000} catalog \citep{Michalik_2015a,Michalik_2015b}.  
After cross-matching all major stellar spectroscopic surveys\footnote{Specifically 
we cross-matched the Tycho-2 catalog against the \project{APOGEE} \citep{Zasowski_2013}, 
\project{GALAH} \citep{DeSilva_2015}, \project{Gaia-ESO} \citep{Gilmore_2012,Randich_2013}, \project{LAMOST}
\citep{Cui_2012}, and \project{RAVE} \citep{Steinmetz_2006} surveys}, we found the RAdial Velocity
Experiment (hereafter \rave) 
survey to have the largest overlap with the first \project{Gaia} data release: 
292,036 stars.  We then used the \project{Gaia} universe model \citep{Robin_2012} to 
estimate the precision in parallax and proper motions that will be available in 
\project{Gaia} DR1 for every star in these cross-matches.  Comparing the 
expected precision to what is currently available, we further found that the 
\project{RAVE} survey will benefit most from \project{Gaia} DR1.  The distances
of 63\% of the \project{RAVE}--\project{Gaia} DR1 overlap sample (182,862 stars)
are expected to improve with the first \project{Gaia} data release, and 47\% of
stars likely to have better proper motions (137,211 stars).  This motivated us to
examine what chemical abundance information is currently available for \rave, and
to evaluate whether we could contribute to this existing set.

@@

\bibitem[Michalik et al.(2015a)]{Michalik_2015a} Michalik, D., Lindegren, L., \& Hobbs, D.\ 2015, \aap, 574, A115

\bibitem[Michalik et al.(2015b)]{Michalik_2015b} Michalik, D., Lindegren, L., Hobbs, D., \& Butkevich, A.~G.\ 2015, \aap, 583, A68
